---
title: STAT210_SP25 Lab 6 Childcare Costs in California
author: River Mckegney
date: today
format: html
---

## The Data

In this lab we're going look at the median weekly cost of childcare in
California. A detailed description of the data can be found [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-05-09/readme.md).

The data come to us from [TidyTuesday](https://github.com/rfordatascience/tidytuesday).

**0. Load the appropriate libraries and the data.**

```{r}
#| label: setup
#| echo: true
#| output: false

library(tidyverse)
```

```{r}
#| label: load-data
#| echo: true
#| output: true
#| eval: true
                                                                                   
options(readr.show_col_types = FALSE)                                                #<1>

childcare_costs <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-09/childcare_costs.csv')

counties <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-09/counties.csv')

#spec(counties) # View col types
#spec(childcare_costs) # View col types
```
1. Source used to hide output warning: <https://github.com/tidyverse/readr/issues/1250>

**1. Briefly describe the dataset (~ 4 sentences). What information does it contain?**

Question 1 Response By River Mckegney:

The 'childcare_costs' dataset contains information about individuals such as unemployment rates at various ages, employment rates by gender and age, income, population size, ethnic background, family size, occupational field, and other demographic data as well as its associated location and date data.
The 'counties' dataset contains state and county data information. Note this data contains location data from various states and not just California.

## California Childcare Costs

Let's start by focusing only on California. 

**2. Create a `ca_childcare` dataset of childcare costs in California, containing (1) county information and (2) just the year and childcare cost variable information from the `childcare_costs` dataset.** 

*Hint: There are 58 counties in CA and 11 years in the dataset. Therefore, your new dataset should have 53 x 11 = 638 observations. The final data set should have study year, median household income expressed in 2018 dollars, all the variables associated with full-time median price charged for Center-based Care, and California county names*

```{r}
#| label: ca-childcare-costs
#| echo: true
#| output: true
#| eval: true

# Merged dataframe
location_childcare_merged <- counties %>% 
  inner_join(y = childcare_costs, by = "county_fips_code")

# Dataframe of selected variables
ca_childcare <- location_childcare_merged %>%
  filter(state_name == "California") %>%
  select(county_name, study_year, mhi_2018, mcsa, mfccsa, mc_infant, mc_toddler, mc_preschool, mfcc_infant, mfcc_toddler, mfcc_preschool) 

```

**3. Using a function from the `forcats` package, complete the code below to create a new variable where each county is categorized into one of the [10 Census regions](https://census.ca.gov/regions/) in California. Use the Region description (from the plot), not the Region number. An example region has been started for you.**  

*Hint: This is probably a good place to use ChatGPT to reduce on tedious work. But you do need to know how to prompt ChatGPT to make it useful!*

```{r}
#| label: recoding-county-to-census-regions
#| echo: true
#| output: true
#| eval: true

ca_childcare <- ca_childcare |>
  mutate(county_name = str_remove(county_name, " County")) |> # Remove the word 'county' from all variable names within the county_name column
  mutate(region = fct_collapse(county_name,                                          # <2>
    'Superior California' = c("Butte", "Colusa", "El Dorado", "Glenn", "Lassen", "Modoc", "Nevada", "Placer", "Plumas", "Sacramento", "Shasta", "Sierra", "Siskiyou", "Sutter", "Tehama", "Yolo", "Yuba"),
    'North Coast' = c("Del Norte", "Humboldt", "Lake", "Mendocino", "Napa", "Sonoma", "Trinity"), 
    'San Francisco Bay Area' = c("Alameda", "Contra Costa", "Marin", "San Francisco", "San Mateo", "Santa Clara", "Solano"),
    'Northern San Joaquin Valley' = c("Alpine", "Amador", "Calaveras", "Madera", "Mariposa", "Merced", "Mono", "San Joaquin", "Stanislaus", "Tuolumne"),
    'Central Coast' = c("Monterey", "San Benito", "San Luis Obispo", "Santa Barbara", "Santa Cruz", "Ventura"),
    'Southern San Joaquin Valley' = c("Fresno", "Inyo", "Kern", "Kings", "Tulare"),
    'Inland Empire' = c("Riverside", "San Bernardino"),
    'Los Angeles County' = c("Los Angeles"),
    'Orange County' = c("Orange"),
    'San Diego - Imperial' = c("Imperial", "San Diego"))) # Create new col 'region' that sorts names into new groupings based on the 10 census region description names

```
2. Reference used to find forcats function required to carry out code script in its given format. <https://forcats.tidyverse.org/reference/index.html>

**4. Let's consider the median household income of each region, and how that income has changed over time. Create a table with ten rows, one for each region, and two columns, one for 2008 and one for 2018. The cells should contain the `median` of the median household income (expressed in 2018 dollars) of the `region` and the `study_year`. Arrange the rows by 2018 values.**

```{r}
#| label: median-income-by-region-over-time
#| eval: true
# Hint: You will want to calculate the median of median income before you make columns for 2008/2018 by transforming the data


```


**5. Which California `region` had the lowest `median` full-time median weekly price for center-based childcare for infants in 2018? Does this `region` correspond to the `region` with the lowest `median` income in 2018 that you found in Q4?**  

*Hint: The code should give me the EXACT answer. This means having the code output the exact row(s) and variable(s) necessary for providing the solution. Consider the `slice()` functions.*

```{r}
#| label: lowest-median-weekly-price-2018
#| eval: true


```

**6. Recreate the plot (look at online instructions) that shows, for all ten regions, the change over time of the full-time median price for center-based childcare for infants, toddlers, and preschoolers. Recreate the plot. You do not have to replicate the exact colors or theme, but your plot should have the same content, including the order of the facets and legend, reader-friendly labels, axes breaks, and a  smooth function.**


```{r}
#| label: recreate-plot
#| eval: true

```


# Data Sources:

1. Data Science Learning Community (2024). Tidy Tuesday: A weekly social data project. https://tidytues.day <https://github.com/rfordatascience/tidytuesday/blob/main/data/2023/2023-05-09/readme.md>

2. United States Department of Labor. Women's Bureau. National Database of Childcare Prices. <https://www.dol.gov/agencies/wb/topics/featured-childcare>

# References:
1. Github. Problem with options(readr.show_col_types = FALSE) #1250. <https://github.com/tidyverse/readr/issues/1250>

2. Wickham H and Posit (2023). Forcats, "Function reference".  <https://forcats.tidyverse.org/reference/index.html>

